---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

### Setup
The objective of this project to create a linear regression model that predicts the popularity of a movie. The given data has been collected from Rotten Tomatoes and IMDb which are popular review-aggregator and database information platforms for films and television.

### Load packages
```{r load-packages, message = FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(gridExtra)
library(GGally)
library(car)
```

### Load data
```{r load-data, echo=TRUE}
load("movies.Rdata")
str(movies)
# Removing all NA rows
movies <- na.omit(movies)
```

## Part 1: Data

The data consists of 651 observations with 32 variables. The data has been randomly sampled from thousands of films and there is no random assignment. This is an observational study so the findings are generalizable but not causal in any way.

## Part 2: Research question

As a person who regularly uses IMDb and Rotten Tomatoes for an experts’ opinion before watching a new movie, I would like to know what makes a movie popular. Being a regular user, I know that ratings and scores are surely good ways to find out how good a movie is but ‘number of votes received’ is a better indicator a movie’s popularity. So is there a relationship between number of votes on IMDb(imdb_num_votes) and the other collected variables like genre, rating, runtime, certification, awards, release time, cast and crew, etc.?

## Part 3: Exploratory data analysis

Exploratory data analysis has been done for selecting valuable features and for better understanding the data. Features showing high variability and interpretability will be selected.

* Features like <mark>title,imdb_url and rt_url</mark> are obviously unique and not useful at all.
```{r Unique-Set,collapse=TRUE}

length(unique(movies$studio))
length(unique(movies$director))
length(unique(movies$actor1))
length(unique(movies$actor2))
length(unique(movies$actor3))
length(unique(movies$actor4))
length(unique(movies$actor5))
```
* All the above features are **high cardinality features** as the number of unique values are close to total number of observations. So, they are not useful for the model.

```{r imdb-titleType and mpaa, echo=TRUE}

p1 <- ggplot(movies, aes(x=title_type, y=imdb_num_votes, fill=title_type)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none")
p2 <- ggplot(movies, aes(x=mpaa_rating, y=imdb_num_votes, fill=mpaa_rating)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none")
grid.arrange(p1, p2, nrow = 1, ncol = 2, top="Distribution of IMDb voter count for different title types and mpaa ratings")
```

* It is not surprising to see that feature films are likely to get much more attention than TV movies and documentaries. Movies rated NC-17 or unrated get unusually low number of votes.

```{r imdb-genre, echo=TRUE}
ggplot(movies, aes(x=genre, y=imdb_num_votes, fill=genre)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none") +
    ggtitle("Distribution of IMDb voter count for different genres") +
    coord_flip()
```

* Action & adventure movies are most likely to get more votes but drama movies have many outliers. Documentaries and art house movies don’t get much attention.

```{r imdb-theatre and DVD release time,fig.height=20, echo=TRUE}
p1 <- ggplot(movies, aes(x=factor(thtr_rel_year), y=imdb_num_votes), fill=factor(thtr_rel_year)) +
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  coord_flip()
p2 <- ggplot(movies, aes(x=factor(dvd_rel_year), y=imdb_num_votes), fill=factor(dvd_rel_year)) +
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  coord_flip()
p3 <- ggplot(movies, aes(x=factor(thtr_rel_month), y=imdb_num_votes), fill=factor(thtr_rel_month)) +
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  coord_flip()
p4 <- ggplot(movies, aes(x=factor(dvd_rel_month), y=imdb_num_votes), fill=factor(dvd_rel_month)) +
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  coord_flip()
p5 <- ggplot(movies, aes(x=factor(thtr_rel_day), y=imdb_num_votes), fill=factor(thtr_rel_day)) +
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  coord_flip()
p6<- ggplot(movies, aes(x=factor(dvd_rel_day), y=imdb_num_votes), fill=factor(dvd_rel_day)) +
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  coord_flip()
grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 3, ncol = 2, top="Distribution of IMDb voter count across DVD and theatre release time")
```

* DVD and theatre releases of recent years are more likely to garner more votes. People don’t seem to vote much for movies released in September, whether released on DVD or in theatres. Aside from thtr_rel_year, none of other features seem to follow any trend or give polarising insights regarding voting.

```{r imdb-Top 200, echo=TRUE}
ggplot(movies, aes(x=top200_box, y=imdb_num_votes, fill=top200_box)) + 
    geom_boxplot(alpha=0.3) +
    ggtitle("Distribution of IMDb voter count for Top 200 highest grossing movies") +
    theme(legend.position="none") 

```

* The striking difference in the distribution of movies based on them being in the top 200 boxofficemojo list indicates movies in this exclusive list are likely to get more votes.

```{r imdb-Oscar wins, echo=TRUE}
p1 <- ggplot(movies, aes(x=best_pic_nom, y=imdb_num_votes, fill=best_pic_nom)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none")
p2 <- ggplot(movies, aes(x=best_pic_win, y=imdb_num_votes, fill=best_pic_win)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none")
p3 <- ggplot(movies, aes(x=best_actor_win, y=imdb_num_votes, fill=best_actor_win)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none")
p4 <- ggplot(movies, aes(x=best_actress_win, y=imdb_num_votes, fill=best_actress_win)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none")
p5 <- ggplot(movies, aes(x=best_dir_win, y=imdb_num_votes, fill=best_dir_win)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none")
grid.arrange(p1, p2, p3, p4, p5, nrow = 1, ncol = 5, top="Distribution of IMDb voter count for different oscar wins")
```

* Films that get nominated or win the oscar are likely to get more votes. Films whose directors are winning oscars are also likely to get more votes.

```{r imdb-Scores and Ratings wins, echo=TRUE}
p1 <- ggplot(movies, aes(x=critics_score, y=imdb_num_votes)) + 
    geom_point()
p2 <- ggplot(movies, aes(x=audience_score, y=imdb_num_votes)) + 
    geom_point()
p3 <- ggplot(movies, aes(x=imdb_rating, y=imdb_num_votes)) + 
    geom_point()
grid.arrange(p1, p2, p3, nrow = 1, ncol = 3, top="Distribution of IMDb voter count for scores and ratings")
```

* Higher approval from audience and critics is likely to result in more votes for a movie.

```{r imdb-Scores CORR, echo=TRUE}
#everything is used for covariances with missing values.
#Pearson is used for indicating corr coefficient
ggcorr(movies, method = c("everything", "pearson"),label=TRUE)
```

* The above correlation heatmap high correlation among audience_score,critics_score and imdb_rating. As this is a collinearity condition, audience_score and critics_score will be dropped in favor imdb_rating. On similar account, dvd_rel_year will be dropped due to high correlation with thtr_rel_year.

## For modeling, the following features will be used:
```{r modelling features, echo=TRUE}
movies2 <- movies %>%
  select(title_type,genre,runtime,mpaa_rating,top200_box,thtr_rel_year,imdb_rating,imdb_num_votes,best_pic_nom,best_pic_win,best_dir_win,thtr_rel_month,thtr_rel_day)
str(movies2)

```

## Part 4: Modeling

For modeling, a multiple regression model with backwards elimination will be used where features will be removed based on high p-values.

First, we will fit all the selected features.
```{r p-values, echo=TRUE}
all_feats <- lm(imdb_num_votes~title_type+genre+runtime+mpaa_rating+top200_box+thtr_rel_year+imdb_rating+best_pic_nom+best_pic_win+best_dir_win,data=movies2)
summary(all_feats)
```
First, we remove <mark>title_type</mark> as all categories have high p-values.

```{r rem title_type}
all_feats_but_1 <- lm(imdb_num_votes~genre+runtime+mpaa_rating+top200_box+thtr_rel_year+imdb_rating+best_pic_nom+best_pic_win+best_dir_win,data=movies2)
summary(all_feats_but_1)
```

Now, we remove <mark>mpaa_rating</mark> as all categories have high p-values.
```{r rem mpaa_rating}
all_feats_but_2 <- lm(imdb_num_votes~genre+runtime+top200_box+thtr_rel_year+imdb_rating+best_pic_nom+best_pic_win+best_dir_win,data=movies2)
summary(all_feats_but_2)
```

Now, we remove <mark>best_dir_win</mark> as it has very high p-value.
```{r rem best_dir_win}
all_feats_but_3 <- lm(imdb_num_votes~genre+runtime+top200_box+thtr_rel_year+imdb_rating+best_pic_nom+best_pic_win,data=movies2)
summary(all_feats_but_3)
```

Some categories of <mark>genre</mark> may have high p-values but some have extremely low p-values too so it should not be removed. Aside from this, all other variable have small p-values so no more removals. This is the final model.

### Model Diagnostics
```{r fig.height=12,fig.width=12}
crPlots(all_feats_but_3)
```

```{r}
plot(all_feats_but_3$residuals ~ movies2$runtime)
abline(0,0)

plot(all_feats_but_3$residuals ~ movies2$thtr_rel_year)
abline(0,0)

plot(all_feats_but_3$residuals ~ movies2$imdb_rating)
abline(0,0)
```

**Linearity condition:** <mark>thtr_rel_year</mark> shows an adequate random scatter around 0. <mark>runtime</mark> and <mark>imdb_rating</mark> on the otherhand, don't exhibit slight or no random scatter of variables, rather concentrated around a specific space. Only <mark>thtr_rel_year</mark> has a strong linear relationship. This might be the reason for the low adjusted R-squared value.

```{r}
hist(all_feats_but_3$residuals,freq=FALSE)
dens <- density(all_feats_but_3$residuals)
lines(dens)
```

There is a slight right skew but the distribution can qualify as nearly normal.
```{r}
qqnorm(all_feats_but_3$residuals)
qqline(all_feats_but_3$residuals)
```

**Nearly normal residuals around mean 0** Slightly significant deviation from mean is towards the tails. The above plots show that the residuals form a nearly normal distribution of plots centered at 0. This condition is met but not strongly.

```{r}
plot(all_feats_but_3$residuals ~ all_feats_but_3$fitted)
abline(0,0)

ncvTest(all_feats_but_3)
```

**Constant variability of residuals:** This condition doesn't seem to be met here as the scatter is not random and is concentrated at a particular space. ncvTest also gives a very significant p-value which means that there is no constant variance.

```{r}
durbinWatsonTest(all_feats_but_3)
```

**Independent residuals:** As p-value form Durbin-Watson test is very high, we can't reject the null hypothesis and conclude that there is no autocorrelation. Hence, these residuals are independent and the condition is met.

### Model interpretation

$\hat{Number.of.votes.on.imdb}$(<mark>imdb_num_votes</mark>) = -5860613.5 +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;190043.1 * <mark>best_pic_win:yes</mark> +<br> 
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;63044.8 * <mark>best_pic_nom:yes</mark> +<br> 
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;38614.7 * <mark>imdb_rating</mark> +<br> 
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;2792.4 * <mark>thtr_rel_year</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;145096.6 * <mark>top200_box:yes</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;1171.8 * <mark>runtime</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;19410.9 * <mark>genre:Science Fiction & Fantasy</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;18778.5 * <mark>genre:Other</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;(-24416.2) * <mark>genre:Mystery & Suspense</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;(-109759.7) * <mark>genre:Musical & Performing Arts</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;(-16328.8) * <mark>genre:Horror</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;(-51219.7) * <mark>genre:Drama</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;(-140098.5) * <mark>genre:Documentary</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;(-15001.0) * <mark>genre:Comedy</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;(-99442.2) * <mark>genre:Art House & International</mark> +<br>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;(-6862.9) * <mark>genre:Animation</mark>

- The intercept is meaningless here as no movie can have value of 0 in multiple features(will lead to extrapolation). In itself, intercept does not provide any useful information but is essential for forming the slope.

- <mark>best_pic_win:yes</mark>: All else held constant, a movie that wins an Oscar for Best Film is likely to be voted by 190043 more users than others.

- <mark>best_pic_nom:yes</mark>: All else held constant, a movie that is nominated an Oscar for Best Film is likely to be voted by 63044 more users than others. 

- <mark>imdb_rating</mark>: All else held constant, around 38614 user votes are needed to increase the imdb rating of a movie by 1.0.

- <mark>thtr_rel_year</mark>: All else held constant, for every one year increase in release, around 2792 more people are likely to vote for a movie.

- <mark>top200_box:yes</mark>: All else held constant, a movie that is in the BoxOfficeMojo Top 200 movies list is likely to be voted by 145096 more users than others.  

- <mark>runtime</mark>: All else held constant, for every one minute increase in runtime, around 1171 more people are likely to vote for a movie.

- <mark>genre</mark>: All else held constant following is the effect of different genres on number of votes
  
  |Genre|Effect|
  |--- |--- |
  |Animation|around 6862 less votes by users|
  |Art House & International|around 99442 less votes by users|
  |Comedy|around 15001 less votes by users|
  |Documentary|around 140098 less votes by users|
  |Drama|around 51219 less votes by users|
  |Horror|around 16328 less votes by users|
  |Musical & Performing Arts|around 109759 less votes by users|
  |Mystery & Suspense|around 24416 less votes by users|
  |Other|around 18778 more votes by users|
  |Science Fiction & Fantasy|around 19410 more votes by users|
  

* * *

## Part 5: Prediction

We will predict the number of votes on IMDb for 21 Jump Street, a comedy film released in 2012.
```{r}
new.data.point <- data.frame(genre="Comedy",runtime=109,top200_box="no",thtr_rel_year=2012,imdb_rating=7.2,best_pic_nom="no",best_pic_win="no")
predict(all_feats_but_3,new.data.point,interval="prediction",level=0.95)
```

We are 95% confident that number of voters on IMDb for 21 Jump Street are between -25394 and 322345. In 2020, 21 Jump Street has around 491000 votes on IMDb. Our prediction interval doesn't contain this value even if take maximum release year in the data into account.


* * *

## Part 6: Conclusion

The model has shown that films with high IMDb rating, winning or getting nominated for a best picture oscar, recent release year, high box office gross(top 200 BoxOfficeMojo) are likely to have more voters i.e., be more popular. More strong features were needed as the standard error of our model i.e., 87860 is very high and R<sup>2</sup>(0.4117) value is also very low. Only 41.17% of variability is explained by the model.

_________________________________________________________________________________________________________________________________________
